Reasoning About a Machine with Local Capabilities

### 2019-02-14 toplas/paper.tex
[TOPLAS] Rewrite StkTokens reference slightly

\section{Discussion}
\label{sec:discussion}
\itoplassug{Reviewer C, esop, would have liked a discussion of (a) what could be done on a machine with a smaller set of capabilities, e.g., no local capability (we could probably not have done anything as the stack pointer could be stored on the heap and reused at unintended times.) and (b) what could be done with a larger/stronger set of available capabilities (we could include a discussion of linear capabilities).}
\itoplassug{Reviewer C, esop, asks whether our attacker model is reasonable (see tex comment). Maybe we should include a short discussion of the attacker model.}
% p.18 "they are linked with an adversary, adv, that is allowed to allocate memory but has no other capabilities".  Is this a reasonable model of the adversary?  (Genuinely naive question.)
\itoplassug{Should we include a discussion of how this ``scales'' to other things like a multi-core setting or if we needed tail calls?}
% * Also, the local capabilities seem related to Boyland's "Borrowing"
%   e.g. that is now supported in Rust amongst other research languages
%   (Pony etc) also mentioned in the ECOOP paper.
% * the combination of E and L capabilities seems very interesting
%   indeed.  I wonder if e.g. they could be used to support strong
%   models of encapsulation e.g. like ownership types. (see eg LNCS
%   7850)
\noindent\textbf{Calling convention}\\
\emph{Formulating control flow correctness} While we claim that our calling convention enforces control-flow correctness, we do not prove a general theorem that shows this because it is not clear what such a theorem should look like.
Formulations in terms of a control-flow graph, like the one by \citet{abadi_control-flow_2005}, do not take into account temporal properties \itoplas{because they only take into account what capabilities can be available and not when they are used.
  Example~\texttt{g1} relies on such temporal properties when returning from the higher-order call backs.}
In fact, our examples show that our logical relation implies a stronger form of control-flow correctness than such formulations, although this is not made very explicit.
In later work, \citet{skorstengaard_stktokens_2019} provide a more explicit and useful definition of control-flow correctness and local state encapsulation.
The idea of their definition is to define a variant of a simple capability machine with a built-in stack as well as call and return instructions.
The call and return instructions use the built-in stack which gives well-bracketed control and local state encapsulation by construction.
They prove full abstraction~\citep{abadi_protection_1998} for a compilation from the capability machine with the built-in stack to the one without which means that the programs running on the capability machine without a built-in stack behaves as though there was a built-in stack.
Their full abstraction proof uses a logical relation, but it differs in a number of ways.
First, their logical relation is a binary cross language logical relation tailored specifically to prove the full abstraction result.
This means that it could not be used to prove program correctness results similar to ours.
Their calling convention is based on something called linear capabilities that offer a different kind of limited revocation.
\itoplassug{Maybe elaborate on why full-abstractness and not weaker properties like ``robust safety''.}

>>>>

\section{Discussion}
\label{sec:discussion}
\itoplassug{Reviewer C, esop, would have liked a discussion of (a) what could be done on a machine with a smaller set of capabilities, e.g., no local capability (we could probably not have done anything as the stack pointer could be stored on the heap and reused at unintended times.) and (b) what could be done with a larger/stronger set of available capabilities (we could include a discussion of linear capabilities).}
\itoplassug{Reviewer C, esop, asks whether our attacker model is reasonable (see tex comment). Maybe we should include a short discussion of the attacker model.}
% p.18 "they are linked with an adversary, adv, that is allowed to allocate memory but has no other capabilities".  Is this a reasonable model of the adversary?  (Genuinely naive question.)
\itoplassug{Should we include a discussion of how this ``scales'' to other things like a multi-core setting or if we needed tail calls?}
% * Also, the local capabilities seem related to Boyland's "Borrowing"
%   e.g. that is now supported in Rust amongst other research languages
%   (Pony etc) also mentioned in the ECOOP paper.
% * the combination of E and L capabilities seems very interesting
%   indeed.  I wonder if e.g. they could be used to support strong
%   models of encapsulation e.g. like ownership types. (see eg LNCS
%   7850)
\noindent\textbf{Calling convention}\\
\emph{Formulating control flow correctness} While we claim that our calling convention enforces control-flow correctness, we do not prove a general theorem that shows this because it is not clear what such a theorem should look like.
Formulations in terms of a control-flow graph, like the one by \citet{abadi_control-flow_2005}, do not take into account temporal properties \itoplas{because they only take into account what capabilities can be available and not when they are used.
  Example~\texttt{g1} relies on such temporal properties when returning from the higher-order call backs.}
In fact, our examples show that our logical relation implies a stronger form of control-flow correctness than such formulations, although this is not made very explicit.
In later work, \citet{skorstengaard_stktokens_2019} provide a more explicit and useful definition of control-flow correctness and local state encapsulation.
The idea of their definition is to define a variant of a simple capability machine with a built-in stack as well as call and return instructions.
The call and return instructions use the built-in stack which gives well-bracketed control and local state encapsulation by construction.
They prove full abstraction~\citep{abadi_protection_1998} for a compilation from the capability machine with the built-in stack to one without which means that the programs running on the capability machine without a built-in stack behaves as though there was a built-in stack.
Their full abstraction proof uses a logical relation, but it differs in a number of ways from the one we present.
Their logical relation is a binary cross language logical relation tailored specifically to prove the full abstraction result.
This means that it could not be used to prove program correctness results similar to ours.
Their calling convention is based on something called linear capabilities rather than local capabilities.
Linear capabilities offer a different kind of limited revocation.
\itoplassug{Maybe elaborate on why full-abstractness and not weaker properties like ``robust safety''.}
